---
title: "Procedure to translate VTT files"
author: "Vladimir Zhbanko"
output:
  html_document:
    df_print: paged
---

### Encrypt your API key in R!

#### Create your personal private and public keys

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE, paged.print=TRUE}
# Generate your private key and write it to the folder, we assume you will save it to the folder C:/Users/UserName/.ssh/ mac users can adapt the path...
# if necessary install package
# install.packages("openssl"); install.packages("tidyverse")
# loads library open ssl and tidyverse
library(openssl)
library(tidyverse)

# private and public key path
path_private_key <- file.path("C:/Users/fxtrams/.ssh", "id_api")
path_public_key <- file.path("C:/Users/fxtrams/.ssh", "id_api.pub")

### KNOW WHAT YOU DOING!!! DO NOT RUN THIS CODE UNNECESARRILY! IT WILL OVERWRITE YOUR KEYS!!!
# generate private key (feel free to customize bits lenght!)
rsa_keygen(bits = 5555) %>% write_pem(path = path_private_key)
# extract and write your public key
read_key(file = path_private_key, password = "") %>% `[[`("pubkey") %>% write_pem(path_public_key)

```

#### Encrypt your API Key

Now you have your personal public key which we will use to encrypt the credentials

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
# encrypt your key (I am showing my API key because I will delete it anyhow after creating the course)
library(openssl)
library(tidyverse)
# private and public key path
path_private_key <- file.path("C:/Users/fxtrams/.ssh", "id_api")
path_public_key <- file.path("C:/Users/fxtrams/.ssh", "id_api.pub")
## Encrypt with your public key - delete the code and clear history files, console
"asdfasfasdfasgadgasfasdfasdfasdfasdf23asdfasdffa" %>% 
  # serialize the object
  serialize(connection = NULL) %>% 
  # encrypt the object
  encrypt_envelope(path_public_key) %>% 
  # write encrypted data to File to your working directory
  write_rds("api_key.enc.rds")

```

Now we have our encrypted key inside our project folder! 

If you will use this script later -> delete API key from your script and feel free to use Version Control Repository!

**NOTE:** if you plan to collaborate and use your key by multiple persons with version control check out R package 'secret'. Remember that you can learn how to use it in my course about Cryptography in R!

### Translate your files!!!

Before translation we should retrieve api_key... We can now return to read our API key:

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
library(openssl)
library(tidyverse)
library(translateR)

# get back our encrypted API key
out <- read_rds("api_key.enc.rds")
# path to our key
path_private_key <- file.path("C:/Users/fxtrams/.ssh", "id_api")
api_key <- decrypt_envelope(out$data, out$iv, out$session, path_private_key, password = "") %>% 
  unserialize()
```

There are a lot of R packages out there... I tried to make another one...

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
# installing my R package from github
devtools::install_github("vzhomeexperiments/translateVTT")

library(translateVTT)

# view package author
citation("translateVTT")
```


```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
# make a list of files to translate
filesToTranslate <-list.files("C:/Users/fxtrams/Downloads/", pattern="*.vtt", full.names=TRUE)
# make a list of languages
languages <- c("fr", "de", "hi", "id","it", "ms", "pt", "ru", "es", "zh-CN", "tr") #add any other language
# get api key
out <- read_rds("api_key.enc.rds")

# decrypting the password using public data list and private key
api_key <- decrypt_envelope(out$data, out$iv, out$session, "C:/Users/fxtrams/.ssh/id_api", password = "") %>% unserialize()
```

# Translation test

```{r}
translateVTT(filesToTranslate[1], sourceLang = "en", destLang = languages[1], apikey = api_key)
```


We are now fully ready to translate but code below would show how to use tryCatch() function to continue for loop in case there is an error with one language

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}

# for loop as more robust function
for (FILE in filesToTranslate) {
  # for loop for languages
  for (LANG in languages) {
    tryCatch({
    # translation
    translateVTT(fileName = FILE, sourceLang = "en", destLang = LANG, apikey = api_key)
    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
  }
}


```

## Results

Your translated files can be found in the same folder as original files...

